---
title: "Mammal sizes"
author: "Philipp Neubauer"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r echo=F,message=F,results='hide'}
require(TaxMeta)
require(knitr)

opts_chunk$set(cache=TRUE)

```



# Variability in mammal size

```{r load insects data}
data('insects')
head(insects)
```  

```?insects``` gives info about the data. First, to apply the hierarchical model, we need to get the taxonomy using the taxize package:

```{r get taxonomy, warning=FALSE, message=FALSE}

taxonomy = c('Species','Genus','Family','Order')

```
We can now apply the meta-analysis:

```{r run anlysis}

insects_ss <- insects %>% filter(!is.na(EADDC),  
                                 !is.na(BDT),
                                 EADDC < 3000,
                                 BDT >-2)

stdise <- function(x) (x-mean(x, na.rm=T))/sd(x, na.rm=T)

insects_ss$lDD <- log10(insects_ss$EADDC)
insects_ss$lUDT <- stdise(insects_ss$UDT)

insects_res_full <- TaxMeta(insects_ss,
                       'lDD',
                       distribution = 'norm',
                       taxonomy = taxonomy,
                       type='full',
                       cont_cov = 'BDT',
                       loo_waic = F,
                       n.chains = 3,
                       n.iter = 50e3,
                       n.thin=25,
                       n.burnin=10e3,
                       return_MCMC=T)
insects_res_full
```

To plot the contributions of different taxonomic levels, simply plot the result:

```{r plot, fig.width = 7,  fig.height = 5}

plot(insects_res_full, 
     taxonomy = taxonomy)

```

To generate predictions at other taxonomic levels, simply add some bogus data to the data. Say, we need an estimate for a species for which the genus was not in the orginal dataset, but the family was part of the dataset:

```{r predictions}

insects_pred <- insects_ss
insects_pred[length(insects_pred)+1,] <- insects_pred[11,]
insects_pred[length(insects_pred)+2,] <- insects_pred[11,]

insects_pred[length(insects_ss)+1,'genus'] <- 'bogus_genus'
insects_pred[length(insects_ss)+1,'species'] <- 'Bogus species'
insects_pred[length(insects_ss)+1,'z_response'] <- NA
insects_pred[length(insects_ss)+2,'species'] <- 'Bogus species'
insects_pred[length(insects_ss)+2,'z_response'] <- NA

insects_res_pred <- TaxMeta(insects_ss,
                       'lDD',
                       distribution = 'norm',
                       taxonomy = taxonomy,
                       type='full',
                       cont_cov = 'BDT',
                       loo_waic = F,
                       n.chains = 3,
                       n.iter = 50e3,
                       n.thin=25,
                       n.burnin=10e3,
                       return_MCMC=F)

insects_res_pred

```

We can also compare alternative model formulations:

```{r run alternative models}

insects_res_fixed <- TaxMeta(insects_ss,
                       'lDD',
                       distribution = 'norm',
                       taxonomy = taxonomy,
                       type='fixed',
                       cont_cov = 'BDT',
                       scale=0.1,
                       loo_waic = T,
                       n.chains = 3,
                       n.iter = 50e3,
                       n.thin=25,
                       n.burnin=10e3,
                       return_MCMC=F)

insects_res_gamma <- TaxMeta(insects_ss,
                       'lDD',
                       distribution = 'norm',
                       taxonomy = taxonomy,
                       type='gamma',
                       cont_cov = 'BDT',
                       loo_waic = T,
                       n.chains = 3,
                       n.iter = 50e3,
                       n.thin=25,
                       n.burnin=10e3,
                       return_MCMC=F)

insects_res_unif <- TaxMeta(insects_ss,
                       'lDD',
                       distribution = 'norm',
                       taxonomy = taxonomy,
                       type='uniform',
                       cont_cov = 'BDT',
                       loo_waic = T,
                       n.chains = 3,
                       n.iter = 50e3,
                       n.thin=25,
                       n.burnin=10e3,
                       return_MCMC=F)

```

And compare the results:

```{r compare}

loo_comp <- loo::compare(insects_res_fixed$loo,
             insects_res_gamma$loo,
             insects_res_unif$loo,
             insects_res_full$loo)

loo_comp <- data.frame(loo_comp)
loo_comp$model <- do.call('rbind',strsplit(rownames(loo_comp),split = '\\$'))[,1]
loo_comp$loo_weight = loo_comp$weight

waic_comp <- loo::compare(insects_res_fixed$waic,
             insects_res_gamma$waic,
             insects_res_unif$waic,
             insects_res_full$waic)

waic_comp <- data.frame(waic_comp)
waic_comp$model <- do.call('rbind',strsplit(rownames(waic_comp),split = '\\$'))[,1]
waic_comp$waic_weight = waic_comp$weight

comp_tab <- dplyr::inner_join(loo_comp,waic_comp,by='model')

knitr::kable(comp_tab[,c('model','looic','se_looic','loo_weight','waic','se_waic','waic_weight')],
             row.names=F,
             digits=c(0,1,1,2,1,1,2),
             align=c('l','r','r','r','r','r','r'))

```

This comparison does not give a good idea of model performance, so lets try a plot:

```{r compare plot, fig.width = 7,  fig.height = 5}

plot(Full=insects_res_full, 
     Fixed =insects_res_fixed, 
     Gamma =insects_res_gamma,
     Uniform=insects_res_unif,
     taxonomy=taxonomy)

```
  
