---
title: "Degree days to maturation for insects"
author: "Philipp Neubauer"
date: "`r Sys.Date()`"
output: 
  pdf_document: 
    latex_engine: xelatex
---

```{r echo=F,message=F,warning=FALSE,results='hide'}
require(TaxMeta)
require(knitr)
require(lme4)

opts_chunk$set(cache=TRUE, autodep=TRUE, warning=FALSE ,message=FALSE)

```



# Taxonomic predictions of degree days to maturation for insects

## Load data

Data is from Jarosik et al, Journal of Entomology - DOI: http://dx.doi.org/10.1603/EC11247.

```{r load insects data}
data('insects')
head(insects)
```  

```?insects``` gives info about the data. First, to apply the hierarchical model, we need to set the taxonomy.

```{r get taxonomy, warning=FALSE, message=FALSE}

taxonomy = c('Species','Genus','Family','Order')

```
## Run taxonomic meta-analysis

After some data manipulation, we can now apply the meta-analysis:

```{r run anlysis, cache=TRUE}


stdise <- function(x) (x-mean(x, na.rm=T))/sd(x, na.rm=T)

insects$lDD <- log10(insects$EADDC)
insects$lBDT <- stdise(insects$BDT)


insects_ss <- insects %>% filter(!is.na(BDT), !is.na(lDD))

lm.ins <- lm(lDD~lBDT+Species,data=insects_ss)
AIC(lm.ins)

lm2.ins <- lm(lDD~lBDT+Species+Genus+Family+Order,data=insects_ss)
AIC(lm2.ins)

lmm.ins <- lmer(lDD~lBDT+(1|Species)+(1|Genus)+(1|Family)+(1|Order),data=insects_ss)
AIC(lmm.ins)

plot(fitted(lm.ins),insects_ss$lDD[!is.na(insects_ss$lDD)])
abline(a=0,b=1)

confint.merMod(lmm.ins,method='Wald')
confint.merMod(lmm.ins)


# reproduce results from 
vcv <- as.data.frame(lme4::VarCorr(lmm.ins))
vcv$vcov/sum(vcv$vcov)
```

These results match those in Jarosik et al. 2011; with most of the variation at the order level (even when including the predictor...).

```{r} 
# out of sample prediction - species
spec <- as.numeric(as.factor(insects_ss$Species))
ssamp <- sample(1:10,max(spec),r=T)
lm.cv.spred = data.frame(matrix(NA,length(spec),3))
for (i in 1:10){
  
  this.lm.ins <- lmer(lDD~lBDT+(1|Species)+(1|Genus)+(1|Family)+(1|Order),data=insects_ss[!spec%in%which(ssamp==i),])
  lm.cv.spred[spec%in%which(ssamp==i),] <-  merTools::predictInterval(this.lm.ins, insects_ss[spec%in%which(ssamp==i),], n.sims=1000, type = "linear.prediction")
  
}

plot(insects_ss$lDD,lm.cv.spred[,1], xlab='Observed', ylab='Predicted')
arrows(insects_ss$lDD, lm.cv.spred[,2], insects_ss$lDD, lm.cv.spred[,3], length=0.05, angle=90, code=3)
abline(a=0,b=1)
```

How does TaxMeta compare?

```{r}

insects_res_full <- TaxMeta(insects[!is.na(insects$lBDT),],
                       'lDD',
                       distribution = 'norm',
                       taxonomy = taxonomy,
                       type='full',
                       cont_cov = 'lBDT',
                       loo_waic = T,
                       n.chains = 3,
                       n.iter = 15e3,
                       n.thin=2,
                       n.burnin=5e3,
                       return_MCMC=T)


lm.pred$est <- 'LME4'
lm.pred$ind <- 1:nrow(lm.pred)

allfx <- summary(insects_res_full$result)[[2]]
tax.pred <- data.frame(allfx[grepl('pred',rownames(allfx)),c(3,1,5)])
colnames(tax.pred) <- colnames(lm.pred)[1:3]
tax.pred$est <- 'TaxMeta'
tax.pred$ind <- 1:nrow(tax.pred)
preds <- rbind(tax.pred,lm.pred)

ggplot(preds) + geom_pointrange(aes(x=ind,
                                    y=fit,
                                    ymin=lwr,
                                    ymax=upr, 
                                    linetype=est),position=position_dodge(width=0.6))

insects_res <- TaxMeta(insects_ss,
                       'lDD',
                       distribution = 'norm',
                       taxonomy = taxonomy,
                       type='full',
                       loo_waic = T,
                       n.chains = 3,
                       n.iter = 50e3,
                       n.thin=25,
                       n.burnin=10e3,
                       return_MCMC=T)
insects_res

plot(insects_ss$lDD,this.cv[,2], pch=16,xlim=c(0,4),ylim=c(0,4))
points(insects_ss$lDD,lm.cv.pred, col=4, pch=16)
arrows(insects_ss$lDD, this.cv[,1], insects_ss$lDD, this.cv[,3], length=0.05, angle=90, code=3)
abline(a=0,b=1)

plot((abs(lm.cv.pred-insects_ss$lDD)),(abs(this.cv[,2]-insects_ss$lDD)),pch=19)
#arrows(insects_ss$lDD, q1, insects_ss$lDD, q3, length=0.05, angle=90, code=3)
abline(a=0,b=1)
df <- data.frame(lmt=(abs(this.cv[,2]-insects_ss$lDD)),lmm=(abs(lm.cv.pred-insects_ss$lDD)))
abline(0,lm(lmt~0+lmm,data=df)$coef,col=2)

this.cvs <- matrix(NA,nrow(insects_ss), 3)
for (i in 1:10){
  insects_test <- insects_ss
  insects_test[spec%in%which(ssamp==i),'lDD'] <- NA
  insects_res_test <- TaxMeta(insects_test,
                              'lDD',
                              distribution = 'norm',
                              taxonomy = taxonomy,
                              type='full',
                              cont_cov = 'lBDT',
                              loo_waic = F,
                              n.chains = 2,
                              n.iter = 50e3,
                              n.thin=25,
                              n.burnin=12e3,
                              return_MCMC=F)
  means <- insects_res_test$result[[1]][,1]
  q1 <- insects_res_test$result[[2]][,1]
  q3 <- insects_res_test$result[[2]][,5]
  pred <- means[grepl('pred',names(means))]
  q1 <- q1[grepl('pred',names(q1))]
  q3 <- q3[grepl('pred',names(q3))]
  this.cvs[spec%in%which(ssamp==i),] <- cbind(q1,pred,q3)
}

par(mfrow=c(1,2))
# LM predictions
plot(insects_ss$lDD,lm.cv.spred[,1], pch=16,xlim=c(1,4),ylim=c(1,4))
arrows(insects_ss$lDD, lm.cv.spred[,2], insects_ss$lDD, lm.cv.spred[,3], length=0.05, angle=90, code=3)
abline(a=0,b=1)


plot(insects_ss$lDD,this.cvs[,2], pch=16,xlim=c(1,4),ylim=c(1,4))
#points(insects_ss$lDD,lm.cv.spred[,1], col=4,pch=16)
#arrows(insects_ss$lDD, lm.cv.spred[,2], insects_ss$lDD, lm.cv.spred[,3], length=0.05, angle=90, code=3,col=4)
arrows(insects_ss$lDD, this.cvs[,1], insects_ss$lDD, this.cvs[,3], length=0.05, angle=90, code=3)
abline(a=0,b=1)

sum(dnorm(insects_ss$lDD,lm.cv.spred[,1],(lm.cv.spred[,3]-lm.cv.spred[,2])/4,log = T))

sum(dnorm(insects_ss$lDD,this.cvs[,2],(this.cvs[,3]-this.cvs[,1])/4,log = T))

plot(abs(lm.cv.spred[,3]-insects_ss$lDD),abs(this.cvs[,3]-insects_ss$lDD))
#arrows(insects_ss$lDD, q1, insects_ss$lDD, q3, length=0.05, angle=90, code=3)
abline(a=0,b=1)

plot(abs(lm.cv.pred-insects_ss$lDD),abs(this.cv[,2]-insects_ss$lDD))

this.cvf <- matrix(NA,nrow(insects_ss), 3)
for (i in 1:10){
  insects_test <- insects_ss
  insects_test[fam%in%which(fsamp==i),'lDD'] <- NA
  insects_res_test <- TaxMeta(insects_test,
                              'lDD',
                              distribution = 'norm',
                              taxonomy = taxonomy,
                              type='full',
                              cont_cov = 'lBDT',
                              loo_waic = F,
                              n.chains = 3,
                              n.iter = 5e3,
                              n.thin=2,
                              n.burnin=2e3,
                              return_MCMC=F)
  means <- insects_res_test$result[[1]][,1]
  q1 <- insects_res_test$result[[2]][,1]
  q3 <- insects_res_test$result[[2]][,5]
  pred <- means[grepl('pred',names(means))]
  q1 <- q1[grepl('pred',names(q1))]
  q3 <- q3[grepl('pred',names(q3))]
  this.cvf[fam%in%which(fsamp==i),] <- cbind(q1,pred,q3)
}

plot(abs(lm.cv.fpred-insects_ss$lDD),abs(this.cvf[,2]-insects_ss$lDD))
plot(lm.cv.fpred,this.cvf[,2])
#arrows(insects_ss$lDD, q1, insects_ss$lDD, q3, length=0.05, angle=90, code=3)
abline(a=0,b=1)


```

To plot the contributions of different taxonomic levels, simply plot the result:

```{r plot, fig.width = 7,  fig.height = 5}

plot(insects_res_full)

plot(Cov = insects_res_full, 
     Taxonomy = insects_res)

```

## Predictions

To generate predictions at other taxonomic levels, simply add some bogus data to the data. Say, we need an estimate for a species for which the genus was not in the orginal dataset, but the family was part of the dataset:

```{r predictions}

insects_pred <- insects_ss
insects_pred[nrow(insects_ss)+1,] <- insects_pred[11,]
insects_pred[nrow(insects_ss)+2,] <- insects_pred[11,]

insects_pred[nrow(insects_ss)+1,'Genus'] <- 'bogus_genus'
insects_pred[nrow(insects_ss)+1,'Species'] <- 'Bogus species'
insects_pred[nrow(insects_ss)+1,'lDD'] <- NA
insects_pred[nrow(insects_ss)+2,'Species'] <- 'Bogus species2'
insects_pred[nrow(insects_ss)+2,'lDD'] <- NA

insects_res_pred <- TaxMeta(insects_pred,
                       'lDD',
                       distribution = 'norm',
                       taxonomy = taxonomy,
                       type='full',
                       cont_cov = 'BDT',
                       loo_waic = F,
                       n.chains = 3,
                       n.iter = 50e3,
                       n.thin=25,
                       n.burnin=10e3,
                       return_MCMC=F)

insects_res_pred

```

## Alternative models
We can also compare alternative model formulations:

```{r run alternative models, cache=TRUE}

insects_res_fixed <- TaxMeta(insects_ss,
                       'lDD',
                       distribution = 'norm',
                       taxonomy = taxonomy,
                       type='fixed',
                       cont_cov = 'BDT',
                       scale=0.01,
                       loo_waic = T,
                       n.chains = 3,
                       n.iter = 5e3,
                       n.thin=2,
                       n.burnin=1e3,
                       return_MCMC=F)

insects_res_gamma <- TaxMeta(insects_ss,
                       'lDD',
                       distribution = 'norm',
                       taxonomy = taxonomy,
                       type='gamma',
                       cont_cov = 'BDT',
                       loo_waic = T,
                       n.chains = 3,
                       n.iter = 50e3,
                       n.thin=25,
                       n.burnin=10e3,
                       return_MCMC=F)

insects_res_unif <- TaxMeta(insects_ss,
                       'lDD',
                       distribution = 'norm',
                       taxonomy = taxonomy,
                       type='uniform',
                       cont_cov = 'BDT',
                       loo_waic = T,
                       n.chains = 3,
                       n.iter = 50e3,
                       n.thin=25,
                       n.burnin=10e3,
                       return_MCMC=F)

```

And compare the results:

```{r compare}

loo_comp <- loo::compare(insects_res_fixed$loo,
             insects_res_gamma$loo,
             insects_res_unif$loo,
             insects_res_full$loo)

loo_comp <- data.frame(loo_comp)
loo_comp$model <- do.call('rbind',strsplit(rownames(loo_comp),split = '\\$'))[,1]
loo_comp$loo_weight = loo_comp$weight

waic_comp <- loo::compare(insects_res_fixed$waic,
             insects_res_gamma$waic,
             insects_res_unif$waic,
             insects_res_full$waic)

waic_comp <- data.frame(waic_comp)
waic_comp$model <- do.call('rbind',strsplit(rownames(waic_comp),split = '\\$'))[,1]
waic_comp$waic_weight = waic_comp$weight

comp_tab <- dplyr::inner_join(loo_comp,waic_comp,by='model')

knitr::kable(comp_tab[,c('model','looic','se_looic','loo_weight','waic','se_waic','waic_weight')],
             row.names=F,
             digits=c(0,1,1,2,1,1,2),
             align=c('l','r','r','r','r','r','r'))

```

This comparison does not give a good idea of model performance, so lets try a plot:

```{r compare plot, fig.width = 7,  fig.height = 5}

plot(Full=insects_res_full, 
     Fixed =insects_res_fixed, 
     Gamma =insects_res_gamma,
     Uniform=insects_res_unif)

```
  
