---
title: "Mammal sizes"
author: "Philipp Neubauer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r echo=F,message=F,results='hide'}
require(TaxMeta)
require(knitr)

opts_chunk$set(cache=TRUE)

```



# Variability in mammal size

```{r load Bergman data, cache=T}
data('sizes')
head(sizes)
```  

```?sizes``` gives info about the data. First, to apply the hierarchical model, we need to get the taxonomy using the taxize package:

```{r get taxonomy, cache=T, warning=FALSE, message=FALSE}

taxonomy = c('species','genus','family','order')

```
We can now apply the meta-analysis:

```{r run anlysis}

species_res_full <- TaxMeta(sizes,
                       'lmass',
                       distribution = 'norm',
                       taxonomy = taxonomy,
                       type='full',
                       loo_waic = F,
                       n.chains = 3,
                       n.iter = 5e3,
                       n.thin=5,
                       n.burnin=1e3,
                       return_MCMC=T)

species_res_full
```

To plot the contributions of different taxonomic levels, simply plot the result:

```{r plot, fig.width = 7,  fig.height = 5}

plot(species_res_full, 
     taxonomy = taxonomy)

```

To generate predictions at other taxonomic levels, simply add some bogus data to the data. Say, we need an estimate for a species for which the genus was not in the orginal dataset, but the family was part of the dataset:

```{r predictions}

Bergman_pred <- Bergman_full
Bergman_pred[length(Bergman_pred)+1,] <- Bergman_pred[11,]
Bergman_pred[length(Bergman_pred)+2,] <- Bergman_pred[11,]

Bergman_pred[length(Bergman_full)+1,'genus'] <- 'bogus_genus'
Bergman_pred[length(Bergman_full)+1,'species'] <- 'Bogus species'
Bergman_pred[length(Bergman_full)+1,'z_response'] <- NA
Bergman_pred[length(Bergman_full)+2,'species'] <- 'Bogus species'
Bergman_pred[length(Bergman_full)+2,'z_response'] <- NA

bergman_res_pred <- TaxMeta(Bergman_pred,
                       'z_response',
                       distribution = 'norm',
                       study_epsilon = 'se',
                       taxonomy = taxonomy,
                       type='full',
                       loo_waic = T,
                       n.chains = 3,
                       n.iter = 150e3,
                       n.thin=30,
                       n.burnin=25e3,
                       return_MCMC=T)

bergman_res_pred

```

We can also compare alternative model formulations:

```{r run alternative models}

bergman_res_fixed <- TaxMeta(Bergman_full,
                       'z_response',
                       distribution = 'norm',
                       study_epsilon = 'se',
                       taxonomy = taxonomy,
                       type='fixed',
                       scale=0.1,
                       loo_waic = T,
                       n.chains = 3,
                       n.iter = 150e3,
                       n.thin=30,
                       n.burnin=25e3,
                       return_MCMC=F)

bergman_res_gamma <- TaxMeta(Bergman_full,
                       'z_response',
                       distribution = 'norm',
                       study_epsilon = 'se',
                       taxonomy = taxonomy,
                       type='gamma',
                       loo_waic = T,
                       n.chains = 3,
                       n.iter = 150e3,
                       n.thin=30,
                       n.burnin=25e3,
                       return_MCMC=F)

bergman_res_unif <- TaxMeta(Bergman_full,
                       'z_response',
                       distribution = 'norm',
                       study_epsilon = 'se',
                       taxonomy = taxonomy,
                       type='uniform',
                       loo_waic = T,
                       n.chains = 3,
                       n.iter = 150e3,
                       n.thin=30,
                       n.burnin=25e3,
                       return_MCMC=F)

```

And compare the results:

```{r compare}

loo_comp <- loo::compare(bergman_res_fixed$loo,
             bergman_res_gamma$loo,
             bergman_res_unif$loo,
             bergman_res_full$loo)

loo_comp <- data.frame(loo_comp)
loo_comp$model <- do.call('rbind',strsplit(rownames(loo_comp),split = '\\$'))[,1]
loo_comp$loo_weight = loo_comp$weight

waic_comp <- loo::compare(bergman_res_fixed$waic,
             bergman_res_gamma$waic,
             bergman_res_unif$waic,
             bergman_res_full$waic)

waic_comp <- data.frame(waic_comp)
waic_comp$model <- do.call('rbind',strsplit(rownames(waic_comp),split = '\\$'))[,1]
waic_comp$waic_weight = waic_comp$weight

comp_tab <- dplyr::inner_join(loo_comp,waic_comp,by='model')

knitr::kable(comp_tab[,c('model','looic','se_looic','loo_weight','waic','se_waic','waic_weight')],
             row.names=F,
             digits=c(0,1,1,2,1,1,2),
             align=c('l','r','r','r','r','r','r'))

```

This comparison does not give a good idea of model performance, so lets try a plot:

```{r compare plot, fig.width = 7,  fig.height = 5}

plot(Full=bergman_res_full, 
     Fixed =bergman_res_fixed, 
     Gamma =bergman_res_gamma,
     Uniform=bergman_res_unif,
     taxonomy=taxonomy)

```
  
